---
title: "Your A4 Report Title (TODO: Revise title)"
author: "Ella Uchtmann"
date: '2022-11-01'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

## Abstract 
[TODO: Include a short abstract (about 50 words)]

## Introduction 
Incarceration in the United States is a huge burden on the shoulders
of racial minorities upheld by the those in power. There are many issues surrounding 
the prison system in the United States. Such as, the disproportionate amount
of black people, specifically black men, incarcerated. In this report, data from 
the Vera Institute of Justice's Incarceration Trends Data Set, will be examined and 
analyzed to further explore inequalities within the United
States prison system. Direct stakeholders that should be considered when researching
this data set are policy makers and government officials. Racism and injustice in
the prison system are upheld and reinforced by policies and laws within the U.S.. 
Change can only truly happen in the hands of well informed people in power through 
the critical analysis and research of this data. Indirect stakeholder that should 
be considered with extra care are people of color, specifically black people. The
decisions made by direct stakeholders based on this data has nearly no daily effect
on those in power. It will drastically impact the lives of thousands of P.O.C. 
people in their everyday lives. 

This report specifically summarizes key values within the data, such as the racial
groups with the largest and smallest populations within the prison system historically
and currently. There are also graphs showing the growth of the prison population
by state and the entirety of the U.S. from 1970 to 2018. The final two charts represent 
inequalities within the U.S. prison system through analysis of the data.  

## Data Summary 

```{r}

```
Through data wrangling I thought it would be insightful to examine the differences
between cumulative data across all years and counties with the most recent year of 
data available. I started with calculating the race with the highest prison population
count across all data. The calculation turned out to be **`r high_prison_count`**. This
means that the racial group with the highest prison population across all data is
Black people. I then decided to see if this trend hold true to the most recent year. 
I calculated the race with the current highest prison population and the result was
**'r current_high_prison_count`**. This shows that Black people continue to hold the 
highest prison population in the U.S.. I then thought it would be useful to understand 
the racial group with the lowest prison counts. I calculated the race the the lowest 
prison population across the whole data set and the result is **`r low_prison_count`**. 
Just as I did with the highest counts, I decided to calculate the race with the current 
lowest prison count to see if this trend still exists. The result is **`r current_low_prison_count`**. 
This means that AAPI peoples hold the lowest prison population historically and currently
as analyzed through this data set. It is important to take into consideration populations
when looking at these results. AAPI and Black peoples make up a relatively small 
individual chunk of the U.S. population. This explains why the AAPI results were
so low. I think this goes to show how powerful these variables truly are. Because, 
while Black people are still a minority group like AAPI people, they still account
for a majority of the prison population despite lesser total population. 

## Growth of the U.S. Prison Population

```{r}

```



## Growth of Prison Population by State

```{r}

```



## TENTATIVE SCATTER PLOT

```{r}

```



## TENTATIVE MAP GRAPH

```{r}

```




