---
title: "Your A4 Report Title (TODO: Revise title)"
author: "Ella Uchtmann"
date: '2022-11-01'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

```{r variable_initialization, echo = FALSE}
bar_chart <- plot_jail_pop_for_us()
line_chart <- plot_jail_pop_by_states()
scatter_plot <- scatter_plot()
map_plot <- map_plot()
```


## Abstract 
[TODO: Include a short abstract (about 50 words)]

## Introduction 
Incarceration in the United States is a huge burden on the shoulders
of racial minorities upheld by the those in power. There are many issues surrounding 
the prison system in the United States. Such as, the disproportionate amount
of black people, specifically black men, incarcerated. In this report, data from 
the Vera Institute of Justice's Incarceration Trends Data Set, will be examined and 
analyzed to further explore inequalities within the United
States prison system. Direct stakeholders that should be considered when researching
this data set are policy makers and government officials. Racism and injustice in
the prison system are upheld and reinforced by policies and laws within the U.S.. 
Change can only truly happen in the hands of well informed people in power through 
the critical analysis and research of this data. Indirect stakeholder that should 
be considered with extra care are people of color, specifically black people. The
decisions made by direct stakeholders based on this data has nearly no daily effect
on those in power. It will drastically impact the lives of thousands of P.O.C. 
people in their everyday lives. 

This report specifically summarizes key values within the data, such as the racial
groups with the largest and smallest populations within the prison system historically
and currently. There are also graphs showing the growth of the prison population
by state and the entirety of the U.S. from 1970 to 2018. The final two charts represent 
inequalities within the U.S. prison system through analysis of the data.  

## Data Summary 

Through data wrangling I thought it would be insightful to examine the differences
between cumulative data across all years and counties with the most recent year of 
data available. I started with calculating the race with the highest prison population
count across all data. The calculation turned out to be **`r high_prison_count`**. This
means that the racial group with the highest prison population across all data is
Black people. I then decided to see if this trend hold true to the most recent year. 
I calculated the race with the current highest prison population and the result was
**'r current_high_prison_count`**. This shows that Black people continue to hold the 
highest prison population in the U.S.. I then thought it would be useful to understand 
the racial group with the lowest prison counts. I calculated the race the the lowest 
prison population across the whole data set and the result is **`r low_prison_count`**. 
Just as I did with the highest counts, I decided to calculate the race with the current 
lowest prison count to see if this trend still exists. The result is **`r current_low_prison_count`**. This means that AAPI peoples hold the lowest prison population historically and currently
as analyzed through this data set. 

It is important to take into consideration populations
when looking at these results. AAPI and Black peoples make up a relatively small 
individual chunk of the U.S. population. This explains why the AAPI results were
so low. I think this goes to show how powerful these variables truly are. Because, 
while Black people are still a minority group like AAPI people, they still account
for a majority of the prison population despite lesser total population. 

## Growth of the U.S. Prison Population

```{r chart1, echo = FALSE}
plot(bar_chart)
```
Bar Chart Showing the Growth of the U.S. Prison Population from 1970 to 2018.

As the chart displays, there seems to be a relatively stagnant prison population
from the early 1970s to the late 1970s at a population just above 150,000. A linear 
increase in the population then began in the late 1970s in to the late 2000s. The
population then settled at about 750,000 people continuing in to 2018. The pattern 
revealed by the chart is that there was a relatively steady low prison population in
the 1970s. Possible changes in policy, societal views, etc. created a shift of hundreds
of thousands of people going into the prison system and then again creating a steady 
prison population but relatively, extremely high.

## Growth of Prison Population by State

```{r chart2, echo = FALSE}
plot(line_chart)
```
Line Chart Showing the Growth of the U.S. Prison Population by States from 1970 to 2018. 

As the chart displays, 

## TENTATIVE SCATTER PLOT

```{r chart3, echo = FALSE}
plot(scatter_plot)
```
Scatter Plot Showing the Difference in Prison Populations fom 


## TENTATIVE MAP GRAPH

```{r chart4, echo = FALSE}
plot(map_plot)
```




